### 2018-03-27 source/egs.tex
added some stuff

>>>>

\begin{table}[h!]
\begin{tabular}{ccccc}
Telescope    & Band   & Resolution  & Area & Reference \\\hline
Chandra      & X-ray  & 0.5''-6.0'' & $17'\times 120'$ & \\
\end{tabular}
\caption{\label{table.LOBOS.sources}Table recapitulating the positions of all 8 chosen calibrator sources, along with their distance from the observation phase centre (EGS phase centre) and calibrator phase centre (3C295), respectively.}
\end{table}

### 2018-03-27 source/intro/RIME.tex
added some stuff

\chapter{The RIME Formalism}
\label{section.RIME}
\pg
While radio interferometry has historically been a complex business (\citepads{2001isra.book.....T}), modern radio interferometry is based on the Radio Interferometer's Measurement Equation, a powerful and elegant formalism which underpins modern calibration and imaging algorithms. In this section, I will give a simplified account of this formalism as described in \citepads{2011A&A...527A.106S} and its companion papers\footnote{\citepads{2011A&A...527A.107S}, \citepads{2011A&A...527A.108S} and \citepads{2011A&A...531A.159S}}. This will form the theoretical basis on which further sections will expand. I cannot overstate the importance of the RIME as the foundation of modern calibration and imaging algorithms.

\pg
For a mathematically rigorous description of the RIME, the references used in this work (particularly \citepads{2011A&A...527A.108S} and \citepads{2011A&A...531A.159S}) are obviously the first place to look. The author takes great care to frame the RIME in the continuity of previous theoretical frameworks of radio interferometry, while also describing it in the context of contemporary software packages and algorithmic tools.

>>>>

\chapter{Introduction to Mathematical Framework: the RIME Formalism}
\pg
In this section, we introduce the mathematical framework which forms the basis for our algorithmic work.
\label{section.RIME}
While radio interferometry has historically been a complex business (\citepads{2001isra.book.....T}), modern radio interferometry is based on the Radio Interferometer's Measurement Equation, a powerful and elegant formalism which underpins modern calibration and imaging algorithms. In this section, I will give a simplified account of this formalism as described in \citepads{2011A&A...527A.106S} and its companion papers\footnote{\citepads{2011A&A...527A.107S}, \citepads{2011A&A...527A.108S} and \citepads{2011A&A...531A.159S}}. This will form the theoretical basis on which further sections will expand. I cannot overstate the importance of the RIME as the foundation of modern calibration and imaging algorithms.

\pg
For a mathematically rigorous description of the RIME, the references used in this work (particularly \citepads{2011A&A...527A.108S} and \citepads{2011A&A...531A.159S}) are obviously the first place to look. We frame the RIME in the continuity of previous theoretical frameworks of radio interferometry, while also describing it in the context of contemporary software packages and algorithmic tools.

### 2018-03-27 source/intro/calibration.tex
added some stuff

\pg
By extending this idea to VLBI\footnote{e.g.  \citepads{1974ApJ...193..293R}}, \emph{amplitude closure} was introduced to the field along with phase closure, as described in \citepads{1983Sci...219...51R}. Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
Third-generation calibration (3GC) is an extension of 2GC calibration which takes direction-dependent effects into account. At the time of writing, this is the state of the art in radio interferometric calibration.

\pg
In this section, I will discuss the approach taken by [cite DDF and kMS paper when it comes out] and [cite prefactor paper if/when it is out], \emph{facet-based calibration}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

\pg

>>>>

\pg
By extending this idea to VLBI\footnote{e.g.  \citepads{1974ApJ...193..293R}}, \emph{amplitude closure} was introduced to the field along with phase closure \citepads[as described in][]{1983Sci...219...51R}. Astronomers were quick to apply these methods to interferometers in general.

@@

\pg
Third-generation calibration (3GC) is an extension of 2GC calibration which takes direction-dependent effects into account. At the time of writing, this is the cutting edge in radio interferometric calibration.

\pg
In this section, I will discuss the approach taken by [cite DDF and kMS paper when it comes out] and [cite prefactor paper if/when it is out], \emph{facet-based calibration}. This method has an imaging equivalent \citepads[see][and associated papers]{2017arXiv171202078T}. The approach cannot be divorced from imaging, for the simple reason that solving for and applying Jones matrices in different directions requires image-plane knowledge of the sky brightness distribution $\Bmatrix$. The imaging challenges this methods introduces will be discussed in [href to relevant imaging section].

\pg
The crucial insight of this approach consists

### 2018-03-27 source/science/3c295.tex
added some stuff

\pg
Having calibrated our 6 subbands with the VLA model, we deconvolve them simultaneously. This allows us to improve the conditioning of the imaging inverse problem. After three passes of self-calibration, we have a spectral cube with 6 frequency slices centred on the various subbands. The stacked residual images are in Fig \ref{fig.3c295.stack.selfcal.uvcut}, before and after self-calibration. We see that noise falls by a factor of [?].
\begin{figure}[h!]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{VLApybdsmModel.FullBW.automate.sc.pass1.app.restored.normalised.fits}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1} Calibrated with the VLA model: there is quite a bit of noise.}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{VLApybdsmModel.FullBW.automate.sc.pass4.app.restored.normalised.fits}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC3} Same data after 3 passes of self-calibration: the noise is decreased.}
\end{subfigure}
\caption{\label{fig.3c295.stack.selfcal.uvcut} Images made with 6 subbands spread across the LOFAR bandwidth.}
\end{figure}
The frequency-dependent model is shown in Fig. [ref]
\pg
               !!! SHOW GAIN CURVES, LOTS OF PLOTS ETC
\pg
               !!! ANALYSE CONVERGENCE OF CALIBRATION ACROSS SUBBANDS
\pg
apply beam to account for different sensitivities of different baselines
\pg
use full-Jones solver (i.e. solve for a 2x2 complex matrix) to account for effects such as Faraday rotation etc
\pg
!!! SHOW JONES CHAIN WE SOLVE FOR AND EXPLAIN IT, MATHS ETC: $\mat{B}\mat{J}$ for Beam*Jones
\pg
!!! SHOW DIRECTION-DEPENDENT PSF EFFECT WHICH WE MODEL

>>>>

\pg
Having calibrated our 6 subbands with the VLA model, we deconvolve them simultaneously. This allows us to improve the conditioning of the imaging inverse problem. After three passes of self-calibration, we have a spectral cube with 6 frequency slices centred on the various subbands. The stacked residual images are in Fig \ref{fig.3c295.stack.selfcal.uvcut}, before and after self-calibration. We see that noise falls by a factor of 1.17. Note that this measure is not as significant as might be expected: indeed, since the scaling factors for each subband is arbitrary, the overall decrease in noise could be attributed to a single subband, should its incorrect scaling factor dominate. The overall decrease is nevertheless encouraging.
\begin{figure}[h!]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{VLApybdsmModel.FullBW.automate.sc.pass1.app.restored.normalised.fits}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1} Calibrated with the VLA model: there is quite a bit of noise.}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{VLApybdsmModel.FullBW.automate.sc.pass4.app.restored.normalised.fits}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC3} Same data after 3 passes of self-calibration: the noise is decreased.}
\end{subfigure}
\caption{\label{fig.3c295.stack.selfcal.uvcut} Images made with 6 subbands spread across the LOFAR bandwidth.}
\end{figure}
The frequency-dependent model is shown in Fig. \ref{fig.3c295.stack.selfcal.uvcut.model}. Note that while there is some frequency-dependent flux distribution changes, it is very small.
\begin{figure}[h!]
\centering
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb035-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb035} SB035}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb095-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb095} SB095}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb095-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb155} SB155}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb095-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb215} SB215}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb095-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb275} SB275}
\end{subfigure}
\hfill
\begin{subfigure}{.43\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{sb095-model}.png}}
\caption{\label{fig.3c295.stack.selfcal.uvcut.SC1.model.sb335} SB335}
\end{subfigure}
\hfill
\caption{\label{fig.3c295.stack.selfcal.uvcut.model} Source model for each subband after 4 passes of self-calibration.}
\end{figure}
\pg
               !!! SHOW GAIN CURVES, LOTS OF PLOTS ETC
\pg
               !!! ANALYSE CONVERGENCE OF CALIBRATION ACROSS SUBBANDS
\pg
apply beam to account for different sensitivities of different baselines
\pg
use full-Jones solver (i.e. solve for a 2x2 complex matrix) to account for effects such as Faraday rotation etc
\pg
!!! SHOW JONES CHAIN WE SOLVE FOR AND EXPLAIN IT, MATHS ETC: $\mat{B}\mat{J}$ for Beam*Jones
\pg
!!! SHOW DIRECTION-DEPENDENT PSF EFFECT WHICH WE MODEL

### 2018-03-27 source/science/hires-EGS.tex
added some stuff

\pg
Show plot of their positions relative to EGS and 3C295

>>>>

\begin{table}[h!]
\begin{tabular}{ccccc}
\# & RA [hms]    & Dec [dms]   & Dist. from EGS [deg.] & Dist. from 3C295 [deg.] \\\hline
1  & 14:30:18.72 & 52:17:29.80 & 2.041                         & 2.904 \\
2  & 14:19:44.44 & 54:23:04.58 & 1.928                         & 2.517 \\ 
3  & 14:21:20.05 & 53:03:46.00 & 0.864                         & 1.743 \\
4  & 14:21:09.41 & 51:22:32.46 & 1.294                         & 1.728 \\
5  & 14:11:50.32 & 52:49:02.66 & 0.844                         & 0.619 \\
6  & 14:11:20.23 & 52:12:04.30 & 0.915                         & 0.000 \\
7  & 14:08:07.00 & 52:55:11.36 & 1.409                         & 0.869 \\
8  & 14:08:09.76 & 52:44:46.56 & 1.354                         & 0.680 \\
\end{tabular}
\caption{\label{table.LOBOS.sources}Table recapitulating the positions of all 8 chosen calibrator sources, along with their distance from the observation phase centre (EGS phase centre) and calibrator phase centre (3C295), respectively.}
\end{table}
\begin{figure}[h!]
\includegraphics[width=0.8\linewidth]{images/EGS_LOBOS_scatterplot}
\caption{Position of the calibrator sources around the EGS. Note that this is not projected properly onto the celestial sphere.}
\label{bootes-coverage-image}
\end{figure}

### 2018-03-27 source/science/strategy.tex
added some stuff

@@

\pg
Observe \& Model 3C295 to get amplitude calibration over entire LOFAR bandwidth

@@

\section{Imaging the EGS with LOFAR international stations}
\pg
if decorrelation is merciful, proceed to patchwise imaging of EGS by using the results from the sections above (DI calibration using 3c295 model, followed by subtraction of all sources seen at low-res except within the patch we want to image; image, change patch; repeat until all EGS imaged)

>>>>

%\pg
%This section aims to provide a general overview of our scientific aims \& chosen strategy. [describe more]

\section{Science Goal}
\pg
One of the key topics this thesis is observational cosmology with LOFAR \& SKA pathfinders. This can be done in many ways - but this manuscript concerns itself primarily with the creation of large, deep and high-resolution surveys of the sky. While this approach is potentially extremely rewarding in terms of the science one can do with its resulting maps, it is also uniquely complicated.

\pg
Why, then, use this approach? One reason is, in fact, for the challenge itself. A high-resolution (matching HST resolution), wide-area (multiple degrees wide), high-quality radio survey has \emph{never} been achieved at the time of writing. Many surveys of the radio sky exist, of course, but they tend to combine only two of those three qualities. FIRST\citepads[see]{2015ApJ...801...26H}, for example, has achieved tremendous results with a $4''$ resolution, relatively low sensitivity, and no short baselines (thereby missing diffuse emission and extended structure). The VLASS \citepads[see]{2013arXiv1312.4602H} will aim to go to a $2.5''$ resolution, medium sensitivity, and few short baselines. %Finally, there will of course be the SKA1-MID survey once the SKA is operational; this survey is expected to provide excellent resolution, sensitivity, and good $uv$-coverage, making it sensitive to diffuse emission as well as point sources.

\pg
In this context, a sky survey made using international LOFAR would be competitive until the SKA-MID survey (and considering it maps the Northern sky rather than the Southern sky, would arguably remain competitive even then). It would also allow LOFAR to fulfill the role of pathfinder in a very literal way, by providing datasets on which algorithms and imaging strategies could be tested before handling SKA data volumes.

\pg
Making a high-resolution full-sky survey using international LOFAR is thus arguably useful in and of itself in terms of interferometric techniques \& instrumentation. What then of its science value? The two most obvious advantages of a good large-sky survey are that they are complete (i.e. they are an accurate sample of the underlying source distribution) while still giving information on rare outlier objects which may happen to lie in the field. This allows scientists a wealth of statistics from which to derive more robust estimations of population distributions and associated scientific measures of interest. Furthermore, depending on the quality of a given survey, it might even be able to provide such statistics simultaneously for various different fields of scientific interest.

\pg
Of course, LOFAR is already providing such catalogues - LOTSS [cite] is only the first of many to come. But these catalogues do not make use of the LOFAR international stations, and thus do not make full use of LOFAR's resolution ($5''$ resolution vs. $0.5''$). Of course, the reason why this is so is because of the technical challenges introduced by the use of international stations. What would their use add to a LOFAR catalogue?

\pg
A high-resolution sky survey has all the advantages listed above, but with the strong benefit of also providing high-resolution images of everyone's favourite objects and sources. It allows for better optical matching and identification, which is particularly relevant e.g. when needing to associate either a low-z or high-z sources to optical counterpart sources (in the case of the Extended Groth Strip, for example, using the international stations means matching the Hubble Space Telescope's resolution). They therefore provide a much more useful contribution to multi-wavelength datasets. It also improves the image's sensitivity to compact sources embedded in diffuse emission: if the source is better-resolved, then the emission associated with compact sources is more easily separated from emission emitted by its surroundings. This can be of tremendous help in image interpretation. Finally, actually resolving objects allows scientists to have better morphological classification for compact objects, which can be extremely useful e.g. when interested in AGN populations vs star-forming galaxy populations.

\pg
talk about AGN science

\pg
talk about star-forming science

@@

\pg
In this section, we briefly describe previous observations of the Extended Groth Strip, an extragalactic field with a rich multi-wavelength coverage described in Table \ref{table.EGS.observation}. It has been long observed as part of the All-Wavelength Extended Groth Strip International Survey collaboration (\citepads{2007ApJ...660L...1D}), which later became part of the CANDELS collaboration (\citepads{2011ApJS..197...35G}). The field is centred at $\alpha=14^h17^m,\delta=+52\deg 30'$, placing it between the tail of Ursa Major and Draco. Its size is $0.7'\times 0.1'$. It has notably been the subject of very deep Hubble Space Telescope observations, which the LOFAR VLBI imaging could attempt to match.

\begin{table}[h!]
\begin{tabular}{cccc}
Telescope    & Band    & Resolution  & Area \\\hline
Chandra      & X-ray   & $0.5''-6.0''$ & $17'\times 120'$ \\
GALEX        & UV      & $5.5''      $ & 1.25$^\text{o}$ diameter \\
HST/ACS      & Optical & $0.1''      $ & $10' \times 67'$\\
HST/NICMOS   & Optical & $0.35''     $ & 0.0128 deg$^2$\\
Megacam      & Optical & $1.0''      $ & 1 deg$^2$\\
IRAC         & IR      & $2.0''      $ & $10' \times 120'$ \\
Spitzer      & IR      & $5.9''-19'' $ & $10'\times 90'$\\
VLA          & Radio   & $1.2''-4.2''$ & $30' \times 80'$
\end{tabular}
\caption{\label{table.EGS.observation}Table recapitulating the multi-wavelength coverage of the Extended Groth Strip, with observations performed as part of the AEGIS (later, CANDELS) collaborations.}
\end{table}

\pg
3C295 is a 3C\footnote{Third Cambridge Catalogue of Radio Sources, a 1959 survey of of the Northern radio sky} source, and thus one of the brighter sources in the Northern radio sky. It lies less than a degree away from the centre of the Extended Groth Strip, which has two important consequences: first, that it can potentially be used as a calibrator source for LOFAR; second, that unless well-modeled and subtracted at the observing instrument's maximum resolution, its sidelobes will pollute the EGS such that very little scientifically useful information might be recovered from a given observation.

@@

\section{Imaging the EGS with LOFAR international stations}

\subsection{Science Goals}

\pg
if decorrelation is merciful, proceed to patchwise imaging of EGS by using the results from the sections above (DI calibration using 3c295 model, followed by subtraction of all sources seen at low-res except within the patch we want to image; image, change patch; repeat until all EGS imaged)
